---
title: "Running AFNI 3dLME"
author: "Dani Cosme"
date: "9/12/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, warning=FALSE, message=FALSE)
```

# Make 3dLME data table
AFNI requires file information be in a [particular format](https://afni.nimh.nih.gov/pub/dist/doc/program_help/3dLME.html). `functional-workshop/code/RX_comparison/AFNI/make_3dLME_dataTable.Rmd` is an example of a script that you could use to put your data into this format.


```{r, child="/Volumes/psych-cog/dsnlab/MDC/functional-workshop/code/RX_comparison/AFNI/make_3dLME_dataTable.Rmd", eval=TRUE}

```

# Make 3dLME bash script
To run the 3dLME model, you will need to create and execute a bash script in which your model is specified.

Path to bash script: `functional-workshop/code/RX_comparison/AFNI/3dLME_all.sh`

```{bash, eval=FALSE}
#!/bin/bash
. ~/.bashrc

# Change directory to the model results folder
cd /Volumes/psych-cog/dsnlab/MDC/functional-workshop/results/AFNI

# Specify 3dLME model
# AFNI documentation https://afni.nimh.nih.gov/pub/dist/doc/program_help/3dLME.html
# Every line in the model must be followed by '\'
#	- prefix = model name
#	- jobs = number of parallel processors
#	- model = model formula
#	- resid = residual file name 
#	- ranEff = random effects, 1 = intercept
#	- SS_type = sum of squares type, 3 = marginal
#	- qVars = quantitative variables
#	- qVars = centering values for quantitative variables
#	- mask = binarized group-level mask
#	- num_glt = number of contrasts (i.e. general linear tests)
#	- gltLabel k = contrast label for contrast k
#	- gltCode k = contrast code for contrast k
#	- datatable = data structure with a header

3dLME -prefix all \
	-jobs 8 \
	-model  "target*domain*age_c+target*domain*age_c2" \
	-resid	all_residuals	\
	-ranEff "~1+age_c" \
	-SS_type 3 \
	-qVars "age_c,age_c2" \
	-qVarCenters "0,0" \
	-mask /Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/RX_mask/groupAverage_opt.nii \
	-num_glt 9 \
	-gltLabel 1 'self-other' -gltCode  1 'target : 1*self -1*other' \
	-gltLabel 2 'social-academic' -gltCode  2 'domain : 1*social -1*academic' \
	-gltLabel 3 'self-other.social-academic' -gltCode  3 'target : 1*self -1*other domain : 1*social -1*academic' \
	-gltLabel 4 'social_self-other' -gltCode  4 'target : 1*self -1*other domain : 1*social' \
	-gltLabel 5 'academic_self-other' -gltCode  5 'target : 1*self -1*other domain : 1*academic' \
	-gltLabel 6 'age.self-other' -gltCode  6 'target : 1*self -1*other age_c : ' \
	-gltLabel 7 'age2.self-other' -gltCode  7 'target : 1*self -1*other age_c2 : ' \
	-gltLabel 8 'age.self-other.social-academic' -gltCode  8 'target : 1*self -1*other domain : 1*social -1*academic age_c : ' \
	-gltLabel 9 'age2.self-other.social-academic' -gltCode  9 'target : 1*self -1*other domain : 1*social -1*academic age_c2 : ' \
	-dataTable \
	Subj	target	domain	age_c	age_c2	InputFile \
	s001	self	academic	-2.58469945	6.6806712468303	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t1_con_0001.nii \
	s001	self	social	-2.58469945	6.6806712468303	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t1_con_0002.nii \
	s001	other	academic	-2.58469945	6.6806712468303	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t1_con_0003.nii \
	s001	other	social	-2.58469945	6.6806712468303	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t1_con_0004.nii \
	s001	self	academic	0.48767123	0.237823228569713	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t2_con_0001.nii \
	s001	self	social	0.48767123	0.237823228569713	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t2_con_0002.nii \
	s001	other	academic	0.48767123	0.237823228569713	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t2_con_0003.nii \
	s001	other	social	0.48767123	0.237823228569713	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t2_con_0004.nii \
	s003	self	academic	-2.46174863	6.06020631730688	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s003_t1_con_0001.nii \
	s003	self	social	-2.46174863	6.06020631730688	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s003_t1_con_0002.nii \
.
.
.
	s096	other	social	-0.12804102	0.0163945028026403	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s096_t2_con_0004.nii \
```

## Specify model
* prefix = model name
* jobs = number of parallel processors
* model = model formula
* ranEff = random effects, 1 = intercept
* SS_type = sum of squares type, 3 = marginal
* qVars = quantitative variables
* qVars = centering values for quantitative variables
* mask = binarized group-level mask
* resid = residual file name 

```{bash, eval=FALSE}
3dLME -prefix all \
	-jobs 8 \
	-model  "target*domain*age_c+target*domain*age_c2" \
	-ranEff "~1+age_c" \
	-SS_type 3 \
	-qVars "age_c,age_c2" \
	-qVarCenters "0,0" \
	-mask /Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/RX_mask/groupAverage_opt.nii \
	-resid	all_residuals	\
```

## Specify contrasts (glts)
* num_glt = number of contrasts (i.e. general linear tests)
* gltLabel k = contrast label for contrast k
* gltCode k = contrast code for contrast k

```{bash, eval=FALSE}
	-num_glt 9 \
	-gltLabel 1 'self-other' -gltCode  1 'target : 1*self -1*other' \
	-gltLabel 2 'social-academic' -gltCode  2 'domain : 1*social -1*academic' \
	-gltLabel 3 'self-other.social-academic' -gltCode  3 'target : 1*self -1*other domain : 1*social -1*academic' \
	-gltLabel 4 'social_self-other' -gltCode  4 'target : 1*self -1*other domain : 1*social' \
	-gltLabel 5 'academic_self-other' -gltCode  5 'target : 1*self -1*other domain : 1*academic' \
	-gltLabel 6 'age.self-other' -gltCode  6 'target : 1*self -1*other age_c : ' \
	-gltLabel 7 'age2.self-other' -gltCode  7 'target : 1*self -1*other age_c2 : ' \
	-gltLabel 8 'age.self-other.social-academic' -gltCode  8 'target : 1*self -1*other domain : 1*social -1*academic age_c : ' \
	-gltLabel 9 'age2.self-other.social-academic' -gltCode  9 'target : 1*self -1*other domain : 1*social -1*academic age_c2 : ' \
```

## Specify data table (input files and design)
* datatable = data structure with a header
* format = subject, condition 1, condition 1, continuous variable 1, continuous variable 2, input file
* Name requirements
  * subject ID column must be named `Subj`
  * file column must be named `InputFile`

```{bash, eval=FALSE}
	-dataTable \
	Subj	target	domain	age_c	age_c2	InputFile \
	s001	self	academic	-2.58469945	6.6806712468303	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t1_con_0001.nii \
	s001	self	social	-2.58469945	6.6806712468303	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t1_con_0002.nii \
	s001	other	academic	-2.58469945	6.6806712468303	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t1_con_0003.nii \
	s001	other	social	-2.58469945	6.6806712468303	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t1_con_0004.nii \
	s001	self	academic	0.48767123	0.237823228569713	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t2_con_0001.nii \
	s001	self	social	0.48767123	0.237823228569713	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t2_con_0002.nii \
	s001	other	academic	0.48767123	0.237823228569713	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t2_con_0003.nii \
	s001	other	social	0.48767123	0.237823228569713	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t2_con_0004.nii \
	s003	self	academic	-2.46174863	6.06020631730688	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s003_t1_con_0001.nii \
	s003	self	social	-2.46174863	6.06020631730688	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s003_t1_con_0002.nii \
.
.
.
	s096	other	social	-0.12804102	0.0163945028026403	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s096_t2_con_0004.nii \
```

## Complete model
```{bash, eval=FALSE}
3dLME -prefix all \
	-jobs 8 \
	-model  "target*domain*age_c+target*domain*age_c2" \
	-resid	all_residuals	\
	-ranEff "~1+age_c" \
	-SS_type 3 \
	-qVars "age_c,age_c2" \
	-qVarCenters "0,0" \
	-mask /Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/RX_mask/groupAverage_opt.nii \
	-num_glt 9 \
	-gltLabel 1 'self-other' -gltCode  1 'target : 1*self -1*other' \
	-gltLabel 2 'social-academic' -gltCode  2 'domain : 1*social -1*academic' \
	-gltLabel 3 'self-other.social-academic' -gltCode  3 'target : 1*self -1*other domain : 1*social -1*academic' \
	-gltLabel 4 'social_self-other' -gltCode  4 'target : 1*self -1*other domain : 1*social' \
	-gltLabel 5 'academic_self-other' -gltCode  5 'target : 1*self -1*other domain : 1*academic' \
	-gltLabel 6 'age.self-other' -gltCode  6 'target : 1*self -1*other age_c : ' \
	-gltLabel 7 'age2.self-other' -gltCode  7 'target : 1*self -1*other age_c2 : ' \
	-gltLabel 8 'age.self-other.social-academic' -gltCode  8 'target : 1*self -1*other domain : 1*social -1*academic age_c : ' \
	-gltLabel 9 'age2.self-other.social-academic' -gltCode  9 'target : 1*self -1*other domain : 1*social -1*academic age_c2 : ' \
	-dataTable \
	Subj	target	domain	age_c	age_c2	InputFile \
	s001	self	academic	-2.58469945	6.6806712468303	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t1_con_0001.nii \
	s001	self	social	-2.58469945	6.6806712468303	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t1_con_0002.nii \
	s001	other	academic	-2.58469945	6.6806712468303	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t1_con_0003.nii \
	s001	other	social	-2.58469945	6.6806712468303	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t1_con_0004.nii \
	s001	self	academic	0.48767123	0.237823228569713	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t2_con_0001.nii \
	s001	self	social	0.48767123	0.237823228569713	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t2_con_0002.nii \
	s001	other	academic	0.48767123	0.237823228569713	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t2_con_0003.nii \
	s001	other	social	0.48767123	0.237823228569713	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s001_t2_con_0004.nii \
	s003	self	academic	-2.46174863	6.06020631730688	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s003_t1_con_0001.nii \
	s003	self	social	-2.46174863	6.06020631730688	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s003_t1_con_0002.nii \
.
.
.
	s096	other	social	-0.12804102	0.0163945028026403	/Volumes/psych-cog/dsnlab/MDC/functional-workshop/data/FX_models/s096_t2_con_0004.nii \
```

# Run 3dLME model
## Dependencies
* AFNI must be installed. Follow the instructions in the [AFNI installation guide](https://afni.nimh.nih.gov/pub/dist/doc/htmldoc/background_install/install_instructs/index.html)
* IF you're running the script locally, make sure the path to AFNI is in your bash environment `~/.bashrc`; if not, export the path to your local version of AFNI using `export PATH=$PATH:/Users/danicosme/AFNI_17.0.12/`
* If you're running it on an HPC clustster, make sure AFNI is loaded in the script; replace `. ~/.bashrc` with `module load afni` or similar
* 3dLME runs using R. To ensure all required R packages are installed, execute this AFNI command: `sudo rPkgsInstall -pkgs ALL`

## Run bash script
1. In the terminal, navigate to the 3dLME bash script and list content
```{bash, eval=TRUE}
cd /Volumes/psych-cog/dsnlab/MDC/functional-workshop/code/RX_comparison/AFNI
ls -l
```

2. 

# Resources
* [AFNI installation guide](https://afni.nimh.nih.gov/pub/dist/doc/htmldoc/background_install/install_instructs/index.html)
* [AFNI 3dLME help guide](https://afni.nimh.nih.gov/pub/dist/doc/program_help/3dLME.html)
* [AFNI message board](https://afni.nimh.nih.gov/afni/community/board/)
